### 2021-02-25 ch07_conclusion/conclusion.tex
Added formatting of key science questions on chapter 07

In a first analysis, we evaluate the climate sensitivity metrics \ac{ECS} and
\ac{TCR} for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\KeyScienceQuestionOne{}}). For \ac{ECS},
we find a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$, which is higher
on the upper and lower end than any model range from previous \ac{CMIP}
generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6 \ac{MMM} of
\ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in \acs{CMIP}6 versus
$3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the upper bound of the
\acs{CMIP}6 models range is considerably larger than in the assessed range of
$\rangeunit{1.5}{4.5}{K}$ given by the latest published \ac{AR} of the
\ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper bound of $4.5
\unit{K}$ is exceeded by about a third of the \acs{CMIP}6 models, with many
models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the model
range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean. For one thing,
this change leads to an improved shortwave \ac{CRE} in these models when
compared to observations. However, this change also substantially reduces the
strong negative shortwave cloud feedback over the Southern Ocean that is
present in previous \ac{CMIP} generations and that results from a cloud phase
change from ice clouds in the present-day to liquid clouds in the future. In
the affected \acs{CMIP}6 models, this cloud phase change is not possible
anymore since the models predominantly simulate liquid clouds over the
present-day Southern Ocean.

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, we evaluate suitable emergent constraints on the \acs{CMIP}6
ensemble. Emergent constraints use a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations. In total we
assess eleven emergent constraints on \ac{ECS}, which are mostly related to
cloud feedbacks since these constitute the most important source of uncertainty
for \ac{ECS} \autocite{Boucher2013, Flato2013}. Since all of the evaluated
emergent constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5
ensemble, we can perform out-of-sample tests on the emergent constraints by
assessing whether they still hold for the \acs{CMIP}6 models. In this study,
which is shown in \cref{ch:05:paper_ecs} and already published in
\textcite{Schlund2020a}, we find a substantial reduction of skill for the
majority of emergent constraints when applied to the \acs{CMIP}6 ensemble in
comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
decrease of the coefficient of determination $R^2$ of the emergent relationship
and a decrease of the statistical significance using the null hypothesis that
there is no correlation between the predictor and \ac{ECS}. Moreover, the
corresponding \acp{PDF} for the emergent constraints show higher $66 \unit{\%}$
\ac{ECS} ranges ($\rangeunit{17}{83}{\%}$ confidence) for almost all emergent
constraints, resulting in values of $\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6
(\acs{CMIP}5: $\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent
constraints, this is an increase in the $66 \unit{\%}$ \ac{ECS} range of $51
\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of
$\rangeunit{2.97}{3.88}{K}$ in \acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in
\acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
emergent constraints. Thus, \cref{enum:01:question_2}
(\enquote{\KeyScienceQuestionTwo{}}) needs to be answered with a \enquote{no}
for the \acs{CMIP}6 ensemble. The increased best estimates and spreads
resulting from the emergent constraints in \acs{CMIP}6 are likely related to
the increased \ac{MMM} and multi-model spread of \ac{ECS} in \acs{CMIP}6. A
possible reason for the reduced skill of the emergent constraints when applied
to the \acs{CMIP}6 ensemble is the increased complexity of the \acs{CMIP}6
models: A basic assumption for these single-process-oriented emergent
constraints is that a single observable process dominates the uncertainty in
\ac{ECS}, which might not be valid anymore due to an increased number of
processes that are included in the \acs{CMIP}6 models.

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. As an alternative, we study projections of
the future \ac{GPP}, which is the largest flux of the terrestrial carbon uptake
and slows down global warming by removing \ac{CO2} from the atmosphere. In the
first step of our new two-step approach, we constrain the global mean \ac{GPP}
at the end of the \nth{21} century in the \acs{CMIP}5 \acs{RCP}8.5 scenario to
$\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint by
\textcite{Wenzel2016}. This first step corrects the \acs{CMIP}5 models' GPP
response to \ac{CO2}, which shows a range of $\rangeunit{156}{248}{\GtCyr{}}$
in the raw multi-model ensemble. In the second step, a \ac{ML}-based climate
model weighting approach is used to further constrain the gridded \ac{GPP}
based on present-day predictors that are relevant for the simulation of
\ac{GPP} in the \acp{ESM}. Our \ac{ML} approach is mathematically similar to
the \ac{MDER} approach \autocite{Karpechko2013, Senftleben2020, Wenzel2016a},
but additionally considers multi-dimensional (gridded) target variables and
non-linear relationships between the predictors and the target variable. We
establish a relationship between process‐oriented predictors and future
projections of \ac{GPP} and then utilize this to project today's observed
conditions into the future. The prediction phase of our method can be
interpreted as an implicit performance weighting. However, due to the complex
structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible to
extract specific values for the individual weights. We consider two target
variables: the gridded monthly climatologies of absolute \ac{GPP}
(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
\nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. We validate our approach by comparing
it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), we find a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) when using our \ac{ML} approach. Moreover, the
evaluation of the global and local feature importance allows us to get further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by our
approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
features. This study directly addresses \cref{enum:01:question_3}
(\KeyScienceQuestionThree{}), which we can answer with \enquote{yes} based on
the results presented here.

>>>>

In a first analysis, we evaluate the climate sensitivity metrics \ac{ECS} and
\ac{TCR} for the latest generation of \acp{ESM} from \acs{CMIP}6. This work,
which is presented in \cref{ch:04:papers_ecs_tcr_assessment} and already
published in \textcite{Bock2020, Meehl2020}, directly addresses
\cref{enum:01:question_1} (\enquote{\emph{\KeyScienceQuestionOne{}}}). For
\ac{ECS}, we find a \acs{CMIP}6 model range of $\rangeunit{1.8}{5.6}{K}$, which
is higher on the upper and lower end than any model range from previous
\ac{CMIP} generations before. In comparison to \acs{CMIP}5, the \acs{CMIP}6
\ac{MMM} of \ac{ECS} is about $16 \unit{\%}$ higher ($3.74 \unit{K}$ in
\acs{CMIP}6 versus $3.23 \unit{K}$ in \acs{CMIP}5). Moreover, especially the
upper bound of the \acs{CMIP}6 models range is considerably larger than in the
assessed range of $\rangeunit{1.5}{4.5}{K}$ given by the latest published
\ac{AR} of the \ac{IPCC} from 2013 \autocite{Stocker2013}. The assessed upper
bound of $4.5 \unit{K}$ is exceeded by about a third of the \acs{CMIP}6 models,
with many models showing \ac{ECS} values above $5 \unit{K}$. For \ac{TCR}, the
model range of \acs{CMIP}6 is $\rangeunit{1.3}{3.0}{K}$, which also exceeds the
\acs{CMIP}5 range of $\rangeunit{1.1}{2.5}{K}$ and the assessed range from
\acs{AR}5 of $\rangeunit{1.0}{2.5}{K}$. A possible reason for the increased
climate sensitivity in many \acs{CMIP}6 models is a change in the microphysical
representation of mixed-phase clouds above the Southern Ocean. For one thing,
this change leads to an improved shortwave \ac{CRE} in these models when
compared to observations. However, this change also substantially reduces the
strong negative shortwave cloud feedback over the Southern Ocean that is
present in previous \ac{CMIP} generations and that results from a cloud phase
change from ice clouds in the present-day to liquid clouds in the future. In
the affected \acs{CMIP}6 models, this cloud phase change is not possible
anymore since the models predominantly simulate liquid clouds over the
present-day Southern Ocean.

In order to reduce this large range of \ac{ECS} in the latest generation of
climate  models, we evaluate suitable emergent constraints on the \acs{CMIP}6
ensemble. Emergent constraints use a physically-based inter-model relationship
between an observable quantity of the Earth system and a target variable to
reduce uncertainties in the target variable with observations. In total we
assess eleven emergent constraints on \ac{ECS}, which are mostly related to
cloud feedbacks since these constitute the most important source of uncertainty
for \ac{ECS} \autocite{Boucher2013, Flato2013}. Since all of the evaluated
emergent constraints have been derived on the \acs{CMIP}3 or \acs{CMIP}5
ensemble, we can perform out-of-sample tests on the emergent constraints by
assessing whether they still hold for the \acs{CMIP}6 models. In this study,
which is shown in \cref{ch:05:paper_ecs} and already published in
\textcite{Schlund2020a}, we find a substantial reduction of skill for the
majority of emergent constraints when applied to the \acs{CMIP}6 ensemble in
comparison to the \acs{CMIP}5 ensemble. This drop in skill is expressed as a
decrease of the coefficient of determination $R^2$ of the emergent relationship
and a decrease of the statistical significance using the null hypothesis that
there is no correlation between the predictor and \ac{ECS}. Moreover, the
corresponding \acp{PDF} for the emergent constraints show higher $66 \unit{\%}$
\ac{ECS} ranges ($\rangeunit{17}{83}{\%}$ confidence) for almost all emergent
constraints, resulting in values of $\rangeunit{1.32}{2.70}{K}$ for \acs{CMIP}6
(\acs{CMIP}5: $\rangeunit{1.16}{1.75}{K}$). Averaged over all emergent
constraints, this is an increase in the $66 \unit{\%}$ \ac{ECS} range of $51
\unit{\%}$. Similarly, the best estimates for \ac{ECS} show values of
$\rangeunit{2.97}{3.88}{K}$ in \acs{CMIP}5 and $\rangeunit{3.48}{4.32}{K}$ in
\acs{CMIP}6, resulting in an increase of about $12 \unit{\%}$ averaged over all
emergent constraints. Thus, \cref{enum:01:question_2}
(\enquote{\emph{\KeyScienceQuestionTwo{}}}) needs to be answered with a
\enquote{no} for the \acs{CMIP}6 ensemble. The increased best estimates and
spreads resulting from the emergent constraints in \acs{CMIP}6 are likely
related to the increased \ac{MMM} and multi-model spread of \ac{ECS} in
\acs{CMIP}6. A possible reason for the reduced skill of the emergent
constraints when applied to the \acs{CMIP}6 ensemble is the increased
complexity of the \acs{CMIP}6 models: A basic assumption for these
single-process-oriented emergent constraints is that a single observable
process dominates the uncertainty in \ac{ECS}, which might not be valid anymore
due to an increased number of processes that are included in the \acs{CMIP}6
models.

To overcome these issues of single-process-oriented emergent constraints,
\cref{ch:06:paper_gpp} introduces an alternative approach based on \ac{ML}.
This work is already published in \textcite{Schlund2020}. Since the new
technique relies on a large number of data points in order to train the \ac{ML}
algorithm, the scalar climate sensitivity expressed as \ac{ECS} or \ac{TCR} is
not an appropriate target variable. As an alternative, we study projections of
the future \ac{GPP}, which is the largest flux of the terrestrial carbon uptake
and slows down global warming by removing \ac{CO2} from the atmosphere. In the
first step of our new two-step approach, we constrain the global mean \ac{GPP}
at the end of the \nth{21} century in the \acs{CMIP}5 \acs{RCP}8.5 scenario to
$\pmrangeunit{171}{12}{\GtCyr{}}$ using a published emergent constraint by
\textcite{Wenzel2016}. This first step corrects the \acs{CMIP}5 models' GPP
response to \ac{CO2}, which shows a range of $\rangeunit{156}{248}{\GtCyr{}}$
in the raw multi-model ensemble. In the second step, a \ac{ML}-based climate
model weighting approach is used to further constrain the gridded \ac{GPP}
based on present-day predictors that are relevant for the simulation of
\ac{GPP} in the \acp{ESM}. Our \ac{ML} approach is mathematically similar to
the \ac{MDER} approach \autocite{Karpechko2013, Senftleben2020, Wenzel2016a},
but additionally considers multi-dimensional (gridded) target variables and
non-linear relationships between the predictors and the target variable. We
establish a relationship between process‐oriented predictors and future
projections of \ac{GPP} and then utilize this to project today's observed
conditions into the future. The prediction phase of our method can be
interpreted as an implicit performance weighting. However, due to the complex
structure of the used \ac{ML} algorithm (\ac{GBRT}) it is not possible to
extract specific values for the individual weights. We consider two target
variables: the gridded monthly climatologies of absolute \ac{GPP}
(\range{2091}{2100}) and the gridded fractional \ac{GPP} change over the
\nth{21} century (2100 versus 2000). The latter quantity shows an increased
\acs{GPP} change in the high latitudes compared to regions closer to the
equator. The results of both approaches are consistent with each other and with
the global constraint of the first step. We validate our approach by comparing
it to other statistical models (the \acs{CMIP}5 \ac{MMM} and a linear
\ac{LASSO} model) in a leave‐one‐model‐out \ac{CV} setup. Compared to \ac{MMM}
(\ac{LASSO}), we find a reduction of the resulting mean \ac{RMSEP} of up to $48
\unit{\%}$ ($3 \unit{\%}$) when using our \ac{ML} approach. Moreover, the
evaluation of the global and local feature importance allows us to get further
insights into the \ac{ML} model. For the first target variable (absolute
\ac{GPP}), historical \ac{GPP} is by far the most important predictor, which
can be explained with a correction of the historical bias in \ac{GPP} by our
approach. For the second variable, \ac{T} and \ac{LAI} are the dominant
features. This study directly addresses \cref{enum:01:question_3}
(\enquote{\emph{\KeyScienceQuestionThree{}}}), which we can answer with
\enquote{yes} based on the results presented here.
